```{r}
setwd("C:\\Users\\panch\\Documents\\Coursework\\Hackathon 2022")
```

#Loading Data
```{r}
data = read.csv("C:\\Users\\panch\\Documents\\Coursework\\Hackathon 2022\\data.csv")
scoringdata = read.csv("C:\\Users\\panch\\Documents\\Coursework\\Hackathon 2022\\scoringData.csv")
```


#Finding na cloumns in data and scoring data
```{r}
na = is.na(data)
sumofna = colSums(is.na(data))
colsumofna = colSums(is.na(data)) > 0
namesofcolswithna = colnames(data)[colSums(is.na(data)) > 0]
namesofcolswithna

na = is.na(scoringdata)
sumofna = colSums(is.na(scoringdata))
colsumofna = colSums(is.na(scoringdata)) > 0
namesofcolswithna = colnames(scoringdata)[colSums(is.na(scoringdata)) > 0]
namesofcolswithna
```

#Replacing NA values with mean of the columns in data and scoringdata
```{r}
data[is.na(data$number_of_ingredients),]$number_of_ingredients <- mean(data$number_of_ingredients, na.rm = T)
scoringdata[is.na(scoringdata$number_of_ingredients),]$number_of_ingredients <- mean(scoringdata$number_of_ingredients, na.rm = T)
```

#Converting categorical variables which are important to factor variables of data and scoringData
```{r}
data$company_location = as.factor(data$company_location)
data$country_of_bean_origin = as.factor(data$country_of_bean_origin)

scoringdata$company_location = as.factor(scoringdata$company_location)
scoringdata$country_of_bean_origin = as.factor(scoringdata$country_of_bean_origin)
```


#Cleaning company location in data and scoringdata
```{r}
library(forcats)
data$company_location = fct_collapse(data$company_location, Europe = c("Belgium","France","Italy","U.K."),
                                                            S.America = c("Ecuador"),
                                                            N.America = c("Canada","U.S.A."))
scoringdata$company_location = fct_collapse(scoringdata$company_location, Europe = c("Belgium","France","Italy","U.K."),
                                                            S.America = c("Ecuador"),
                                                            N.America = c("Canada","U.S.A."))

```

#Forming Binary Variables from Boolean Variable Columns in data and scoringdata
```{r}
data$sweet_binary <- ifelse(data$sweet == 'Yes', 1, 0)
data$nutty_binary <- ifelse(data$nutty == 'Yes', 1, 0)
data$cocoa_binary <- ifelse(data$cocoa == 'Yes', 1, 0)
data$roasty_binary <- ifelse(data$roasty == 'Yes', 1, 0)
data$earthy_binary <- ifelse(data$earthy == 'Yes', 1, 0)
data$creamy_binary <- ifelse(data$creamy == 'Yes', 1, 0)
data$sandy_binary <- ifelse(data$sandy == 'Yes', 1, 0)
data$fatty_binary <- ifelse(data$fatty == 'Yes', 1, 0)
data$floral_binary <- ifelse(data$floral == 'Yes', 1, 0)
data$intense_binary <- ifelse(data$intense == 'Yes', 1, 0)
data$sour_binary <- ifelse(data$sour == 'Yes', 1, 0)
data$vanilla_binary <- ifelse(data$vanilla == 'Yes', 1, 0)
data$fruit_binary <- ifelse(data$fruit == 'Yes', 1, 0)
data$molasses_binary <- ifelse(data$molasses == 'Yes', 1, 0)
data$woody_binary <- ifelse(data$woody == 'Yes', 1, 0)
data$sticky_binary <- ifelse(data$sticky == 'Yes', 1, 0)
data$coffee_binary <- ifelse(data$coffee == 'Yes', 1, 0)
data$rich_binary <- ifelse(data$rich == 'Yes', 1, 0)
data$dried.fruit_binary <- ifelse(data$dried.fruit == 'Yes', 1, 0)
data$ingredient_Beans_binary <- ifelse(data$ingredient_Beans == 'Yes', 1, 0)
data$ingredient_Sugar_binary <- ifelse(data$ingredient_Sugar == 'Yes', 1, 0)
data$ingredient_Sweetener_binary <- ifelse(data$ingredient_Sweetener == 'Yes', 1, 0)
data$ingredient_Cocoa_Butter_binary <- ifelse(data$ingredient_Cocoa_Butter == 'Yes', 1, 0)
data$ingredient_vanilla_binary <- ifelse(data$ingredient_vanilla == 'Yes', 1, 0)
data$ingredient_lecithin_binary <- ifelse(data$ingredient_lecithin == 'Yes', 1, 0)
data$ingredient_salt <- ifelse(data$ingredient_salt == 'Yes', 1, 0)


scoringdata$sweet_binary <- ifelse(scoringdata$sweet == 'Yes', 1, 0)
scoringdata$nutty_binary <- ifelse(scoringdata$nutty == 'Yes', 1, 0)
scoringdata$cocoa_binary <- ifelse(scoringdata$cocoa == 'Yes', 1, 0)
scoringdata$roasty_binary <- ifelse(scoringdata$roasty == 'Yes', 1, 0)
scoringdata$earthy_binary <- ifelse(scoringdata$earthy == 'Yes', 1, 0)
scoringdata$creamy_binary <- ifelse(scoringdata$creamy == 'Yes', 1, 0)
scoringdata$sandy_binary <- ifelse(scoringdata$sandy == 'Yes', 1, 0)
scoringdata$fatty_binary <- ifelse(scoringdata$fatty == 'Yes', 1, 0)
scoringdata$floral_binary <- ifelse(scoringdata$floral == 'Yes', 1, 0)
scoringdata$intense_binary <- ifelse(scoringdata$intense == 'Yes', 1, 0)
scoringdata$sour_binary <- ifelse(scoringdata$sour == 'Yes', 1, 0)
scoringdata$vanilla_binary <- ifelse(scoringdata$vanilla == 'Yes', 1, 0)
scoringdata$fruit_binary <- ifelse(scoringdata$fruit == 'Yes', 1, 0)
scoringdata$molasses_binary <- ifelse(scoringdata$molasses == 'Yes', 1, 0)
scoringdata$woody_binary <- ifelse(scoringdata$woody == 'Yes', 1, 0)
scoringdata$sticky_binary <- ifelse(scoringdata$sticky == 'Yes', 1, 0)
scoringdata$coffee_binary <- ifelse(scoringdata$coffee == 'Yes', 1, 0)
scoringdata$rich_binary <- ifelse(scoringdata$rich == 'Yes', 1, 0)
scoringdata$dried.fruit_binary <- ifelse(scoringdata$dried.fruit == 'Yes', 1, 0)
scoringdata$ingredient_Beans_binary <- ifelse(scoringdata$ingredient_Beans == 'Yes', 1, 0)
scoringdata$ingredient_Sugar_binary <- ifelse(scoringdata$ingredient_Sugar == 'Yes', 1, 0)
scoringdata$ingredient_Sweetener_binary <- ifelse(scoringdata$ingredient_Sweetener == 'Yes', 1, 0)
scoringdata$ingredient_Cocoa_Butter_binary <- ifelse(scoringdata$ingredient_Cocoa_Butter == 'Yes', 1, 0)
scoringdata$ingredient_vanilla_binary <- ifelse(scoringdata$ingredient_vanilla == 'Yes', 1, 0)
scoringdata$ingredient_lecithin_binary <- ifelse(scoringdata$ingredient_lecithin == 'Yes', 1, 0)
scoringdata$ingredient_salt <- ifelse(scoringdata$ingredient_salt == 'Yes', 1, 0)



```

#Finding correlation of all numeric variables with price
```{r}
numericVars <- which(sapply(data, is.numeric))
data_numVar <- data[,numericVars]
cor_numVar <- cor(data_numVar)
cor_sorted <- as.matrix(sort(cor_numVar[,'rating'], decreasing = TRUE))

(cor_sorted)
```

#Exporting correalation matrix image for presentation 
```{r}
#png(filename = "mycorrplot.png", width = 1200, height = 800)
#corrplot(cor_numVar, type = "lower",)

```


#Stating the column types in the data
```{r}
NumericVars <- which(sapply(data, is.numeric)) #index vector numeric variables
FactorVars <- which(sapply(data, is.factor)) #index vector factor variables
CharacterVars <- which(sapply(data, is.character)) #index vector character variables

cat('There are', length(NumericVars), 'numeric variables, and', length(FactorVars), 'factor variables, and',length(CharacterVars),'character variables')
```
#Subsetting data and scoring data to make models
```{r}
DATA = subset(data ,select = c(rating,cocoa_binary,creamy_binary,ingredient_Beans_binary,rich_binary,fruit_binary,earthy_binary,number_of_ingredients,sweet_binary,sour_binary,cocoa_percent,ingredient_vanilla_binary,vanilla_binary,company_location,dried.fruit_binary,nutty_binary,fatty_binary,molasses_binary))
              
SCORINGDATA = subset(scoringdata,select = c(cocoa_binary,creamy_binary,ingredient_Beans_binary,rich_binary,fruit_binary,earthy_binary,number_of_ingredients,sweet_binary,sour_binary,cocoa_percent,ingredient_vanilla_binary,vanilla_binary,company_location,dried.fruit_binary,nutty_binary,fatty_binary,molasses_binary))

SCORINGDATA$rating = NA
```

#Splitting the data into test and train
```{r}
library(caret)
set.seed(617)
split<-createDataPartition(DATA$rating,p = 0.75 ,list=FALSE,groups = 100)
train<-DATA[split,]
test<-DATA[-split,]

```

#Best linear regression
```{r}
model_lm = lm(rating ~ . , train)

pred_lm_train = predict(model_lm)
rmse_lm_train = sqrt(mean((pred_lm_train - train$rating)^2))

pred_lm = predict(model_lm , newdata = test)
rmse_lm_test = sqrt(mean((pred_lm - test$rating)^2))

 

rmse_lm_train ; rmse_lm_test
```

#Best logistic regression
```{r}

model_glm = glm(rating ~ ., train , family = 'inverse.gaussian')

pred_glm_train = predict(model_glm, type='response')
rmse_glm_train = sqrt(mean((pred_glm_train - train$rating)^2)) #70.82915

pred_glm_test = predict(model_glm , newdata = test , type = 'response')
rmse_glm_test = sqrt(mean((pred_glm_test - test$rating)^2)) #71.19982

#pred_glm_SCORING = predict(model_lm , newdata = SCORINGDATA , type = 'response')

rmse_glm_train ; rmse_glm_test

#(test rmse : 70.13889 , rmse after uploading : 69.96191)
```

#Best decision tree selection
```{r}

library(rpart)

model_tree = rpart(rating ~ . , train , method ='anova', control = rpart.control(cp = 4e-04))

pred_tree_train = predict(model_tree)
rmse_tree_train = sqrt(mean((pred_tree_train - train$rating)^2)) #65.26914

pred_tree_test = predict(model_tree , newdata = test)
rmse_tree_test = sqrt(mean((pred_tree_test - test$rating)^2)) #68.48148

#pred_tree_SCORING = predict(model_tree , newdata = SCORINGDATA)

rmse_tree_train ; rmse_tree_test
```

#Best Ranger model
```{r}
library(ranger)
set.seed(617)
ranger = ranger(rating~.,
                       data = train, 
                       num.trees = 3000)

pred_ranger_train = predict(ranger, data = train, num.trees = 3000)
rmse_ranger_train = sqrt(mean((pred_ranger_train$predictions - train$rating)^2))


pred_ranger_test = predict(ranger , test , num.trees = 3000)
rmse_ranger_test = sqrt(mean((pred_ranger_test$predictions - test$rating)^2))


rmse_ranger_train ; rmse_ranger_test
```

#Preparing data in a matrix format to feed in the XGboost model
```{r}
library(vtreat)
trt = designTreatmentsZ(dframe = train,
                        varlist = names(train)[2:18])

newvars = trt$scoreFrame[trt$scoreFrame$code%in% c('clean','lev'),'varName']

train_input = prepare(treatmentplan = trt, 
                      dframe = train,
                      varRestriction = newvars)

test_input = prepare(treatmentplan = trt, 
                     dframe = test,
                     varRestriction = newvars)

SCORING_input = prepare(treatmentplan = trt, 
                     dframe = SCORINGDATA,
                     varRestriction = newvars)
```

#Best XGboost model
```{r}
library(xgboost)
set.seed(617)
xgboost12 <- xgboost(data = as.matrix(train_input), 
               label = train$rating, 
               eta = 0.001,gamma=0,
               max_depth = 3,
               nround=6000, 
               objective="reg:squarederror",
               early_stopping_rounds=50,
               print_every_n=1000,
               subsample = 0.5,
               )


pred_xgb12_train = predict(xgboost12 , newdata = as.matrix(train_input))
rmse_xgb12_train = sqrt(mean((pred_xgb12_train - train$rating)^2))

pred_xgb12_test = predict(xgboost12 , newdata = as.matrix(test_input))
rmse_xgb12_test = sqrt(mean((pred_xgb12_test - test$rating)^2))

rmse_xgb12_train ; rmse_xgb12_test
```

Best Gradient Boosting Model
```{r}
library(gbm)
set.seed(617)
boost = gbm(rating~.,
            data=train,
            distribution="gaussian",
            n.trees = 3000,
            interaction.depth = 2,
            shrinkage = 0.01)


pred_gbm_train = predict(boost , train )
rmse_gbm_train = sqrt(mean((pred_gbm_train - train$rating)^2))

pred_gbm_test = predict(boost , newdata = test )
rmse_gbm_test = sqrt(mean((pred_gbm_test - test$rating)^2))


pred_gbm_SCORING = predict(boost , newdata = SCORINGDATA)

rmse_gbm_train ; rmse_gbm_test
```
#Displaying the results of the Gradient Boosting Model
```{r}
par(mar = c(7, 11, 1, 1))
summary(
  boost, 
  cBars = 15,
  method = relative.influence, # also can use permutation.test.gbm
  las = 2
  )
```
#Finding the relationship of cocoa percent and rating of chocolate
```{r}
model = lm(rating ~ cocoa_percent,data)
summary(model)
```

#Finding the relationship of company location and rating of chocolate
```{r}
model = lm(rating ~ relevel(company_location,ref='Other'),data)
summary(model)
```

#Finding the relationship of number_of_ingredients and rating of chocolate
```{r}
model = lm(rating ~ number_of_ingredients,data)
summary(model)
```

#Making predictions and forming the submission file
```{r}
pred_gbm_SCORING = predict(boost , newdata = SCORINGDATA)

id = scoringdata$id
predicted_rating = pred_gbm_SCORING
file = data.frame(id , predicted_rating)      

sub1 = write.csv(x = file , file = 'TeamADA.csv' , row.names = F)
```

